Automated Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges

### 2018-03-08 paper.tex
fixed typo
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
updated title
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
mods to section 2

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Automated Program Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}

@@

\author{Eddie B. de Lima Filho}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{TP Vision}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

@@

Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.

@@

The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they should be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is the stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region. The stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.

Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest studies on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems. In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing. Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.

Along with CPS modeling and verification techniques purely based on HA, one can also highlight logic-based approaches~\cite{Phan15} that may still use HA, but are not based on a algebraic description of states and transitions. Some examples of logic-based verification are theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among which BMC stands out~\cite{handbook09}.

%Much of the work related to the formal verification of ECPS seeks to ensure the safety, robustness, and reachability properties of an HA. 
Recently, some studies have addressed ECPS formal verification considering discretization, switching, and quantization effects.  Dugirala and Viswanathan~\cite{Duggirala2015} verified embedded control software considering real time issues, {\it e.g.}, delayed responses and possible loss of real time deadlines. Anta {\it et al.}~\cite{Anta10} presented a tool called Costan, which finds errors in mathematical-model implementations and verifies whether error is tolerated, considering quantization effects and fixed-point implementation, while focusing on quantization error effects over system stability. Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} preseted DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends, in order to provide support for digital system design and verification considering FWL effects. In particular, DSVerifier checks if a designed digital controller and/or closed-loop ECPS presents the desired performance, when it is implemented with a FWL format~\cite{Bessa16,Bessa17}.

Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}. Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems, which are hardly modeled through automata- or logig-based approaches. Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs. In order to achieve a balanced trade-off between ECPS-simulation scalability and correctness of verification methods, statistical model checking have been employed to formally verify ECPSs~\cite{Clarke11}.

@@

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification}
\label{fig:CEGIS-precision-increment}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Automated Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}

@@

\author{Eddie B. de Lima Filho}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{TPV Technology}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

@@

Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as provide controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.

@@

\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho... parece que estamos misturando tudo... sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo muito nebuloso?}
The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled, and they should be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region. Stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification} or linear system stability theory for discrete simulations~\cite{Bessa17}.

Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest studies on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems. In the last decade, signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing. Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.

Along with CPS modeling and verification techniques purely based on HA, one can also highlight logic-based approaches~\cite{Phan15} that may still use HA, but which are not based on a algebraic description of states and transitions. Some examples of logic-based verification are theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among which BMC stands out~\cite{handbook09}.

%Much of the work related to the formal verification of ECPS seeks to ensure the safety, robustness, and reachability properties of an HA. 
Recently, some studies have addressed ECPS formal verification considering discretization, switching, and quantization effects.  Dugirala and Viswanathan~\cite{Duggirala2015} verified embedded control software considering real time issues, {\it e.g.}, delayed responses and possible loss of real time deadlines. Anta {\it et al.}~\cite{Anta10} presented a tool called Costan, which finds errors in mathematical-model implementations and verifies whether error is tolerated, considering quantization effects and fixed-point implementation, while focusing on quantization error effects over system stability. Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} preseted DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends, in order to provide support for digital system design and verification considering FWL effects. In particular, DSVerifier checks if a designed digital controller and/or closed-loop ECPS presents the desired performance, when it is implemented with a given FWL format~\cite{Bessa16,Bessa17}.

Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}. Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems that are hardly modeled through automata- or logig-based approaches. Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs. In order to achieve a balanced trade-off between ECPS-simulation scalability and correctness of verification methods, statistical model checking have been employed to formally verify ECPSs~\cite{Clarke11}.

@@

 \end{tikzpicture}
}}
\caption{\textcolor{blue}{CEGIS with multi-staged verification for ECPS.}}
\label{fig:CEGIS-precision-increment}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
